Shift-sensitive Power Transformations for Transforming Data to Normality

### 2023-11-01 manuscript/manuscript.Rmd
Fixes to manuscript.

```{=latex}
\begin{equation}
\phi_{\text{BC}}^\lambda (x_i) = 
\begin{cases}
\left(x_i^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
\log(x_i) & \text{if } \lambda = 0
\end{cases}
\end{equation}
```
One limitation of th Box-Cox transformation is that it is only defined for $x_i > 0$. In contrast, the Yeo-Johnson transformation under the transformation parameter $\lambda$ is defined for any $x_i \in \mathbb{R}$:

@@

```{r reduced-normality, echo=FALSE, fig.cap=cap, warning=FALSE}
cap <- paste0(
  "Power transformations may reduce normality depending on location. ",
  "Variable $x$ consists of $10000$ values are randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$. ",
  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
  "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
  "and even breaks down due to numerical issues for $\\mu > 5 10^4$.")

@@

```{=latex}
\begin{equation}
\left\{ \hat{\lambda}^{\text{rBC}}, \hat{x}_0^{\text{rBC}} \right\} = \argmax_{\lambda, x_0} \frac{1}{\sum_{i=1}^n w_u} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2_w) + (\lambda - 1) \log(x_i - x_0) \right]
\end{equation}
```
where $\sigma^2_w$ is the weighted variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0} (\mathbf{X})$:

@@

```{=latex}
\begin{equation}
\left\{ \hat{\lambda}^{\text{rYJ}}, \hat{x}_0^{\text{rYJ}} \right\} = \argmax_{\lambda, x_0} \frac{1}{\sum_{i=1}^n w_u} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2_w) + (\lambda - 1) \sgn(x_i - x_0) \log(1 + |x_i - x_0|) \right]
\end{equation}
```
where $\sigma^2_w$ is the weighted variance of the Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0} (\mathbf{X})$:

@@

- Set weights based on z-score of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$. After @Raymaekers2021-kq, $z_i = \frac{\phi^{\lambda, x_0}(x_i) - \mu_M}{\sigma_M}$. Here, $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$ [@Huber1981-su].

@@

All weighting functions share the characteristic that for $\dot{x}_i < k_1$, instances are fully weighted, i.e. when $k_1 > 0$ the weighting functions are symmetric window functions with a flat top. The triangle and tapered cosine functions then gradually down-weight instances with $k_1 \leq \dot{x}_i \leq k_2$, and assign no weight to instances $x_i > k_2$.

@@

.plot_weighting_functions(plot_theme = plot_theme)
```
Other criteria aside from MLE can be used for finding robust transformation parameters. We define robust criteria in Appendix A **insert**.

@@

The Cramér-von Mises criterion is also based on the empirical distribution function of $\mathbf{X}$. We define the Cramér-von Mises criterion as follows:
```{=latex}
\begin{equation}
U_{\text{CvM}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \right)^2}
\end{equation}
```

>>>>

```{=latex}
\begin{equation}
\phi_{\text{BC}}^\lambda (x_i) = 
\begin{cases}
\left(x_i^\lambda - 1 \right) / \lambda & \text{if } \lambda \neq 0\\
\log(x_i) & \text{if } \lambda = 0
\end{cases}
\end{equation}
```
One limitation of the Box-Cox transformation is that it is only defined for $x_i > 0$. In contrast, the Yeo-Johnson transformation under the transformation parameter $\lambda$ is defined for any $x_i \in \mathbb{R}$:

@@

```{r reduced-normality, echo=FALSE, fig.cap=cap, warning=FALSE}
cap <- paste0(
  "Power transformations may reduce normality depending on location. ",
  "Variable $x$ consists of $10000$ values are randomly drawn from a normal distribution $\\mathcal{N}(\\mu, 1)$. ",
  "For both Box-Cox and Yeo-Johnson transformations, we would expect $\\lambda = 1$, since $x$ is already normally distributed. ",
  "The figure shows that instead, $\\lambda$ is strongly dependent on the location $\\mu$, ",
  "and even breaks down due to numerical issues for $\\mu > 5 \cdot 10^4$.")

@@

```{=latex}
\begin{equation}
\left\{ \hat{\lambda}^{\text{rBC}}, \hat{x}_0^{\text{rBC}} \right\} = \argmax_{\lambda, x_0} \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2_w) + (\lambda - 1) \log(x_i - x_0) \right]
\end{equation}
```
where $\sigma^2_w$ is the weighted variance of the Box-Cox transformed feature $\phi_{\text{BC}}^{\lambda, x_0} (\mathbf{X})$:

@@

```{=latex}
\begin{equation}
\left\{ \hat{\lambda}^{\text{rYJ}}, \hat{x}_0^{\text{rYJ}} \right\} = \argmax_{\lambda, x_0} \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left[ -\frac{1}{2}\log(\sigma^2_w) + (\lambda - 1) \sgn(x_i - x_0) \log(1 + |x_i - x_0|) \right]
\end{equation}
```
where $\sigma^2_w$ is the weighted variance of the Yeo-Johnson transformed feature $\phi_{\text{YJ}}^{\lambda, x_0} (\mathbf{X})$:

@@

- Set weights based on the z-score of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$. After @Raymaekers2021-kq, $z_i = \frac{\phi^{\lambda, x_0}(x_i) - \mu_M}{\sigma_M}$. Here, $\mu_M$ and $\sigma_M$ are robust Huber M-estimates of location and scale of the transformed feature $\phi^{\lambda, x_0} (\mathbf{X})$ [@Huber1981-su].

@@

All weighting functions share the characteristic that for $\left| \dot{x}_i \right|< k_1$, instances are fully weighted, i.e. when $k_1 > 0$ the weighting functions are symmetric window functions with a flat top. The triangle and tapered cosine functions then gradually down-weight instances with $k_1 \leq \left| \dot{x}_i \right| \leq k_2$, and assign no weight to instances $\left| x_i \right| > k_2$.

@@

.plot_weighting_functions(plot_theme = plot_theme)
```
Other criteria aside from MLE can be used for finding robust transformation parameters. We define robust criteria in Appendix A **INSERT**.

@@

The Cramér-von Mises criterion is also based on the empirical distribution function of $\mathbf{X}$. We define the Cramér-von Mises criterion as follows:
```{=latex}
\begin{equation}
U_{\text{CvM}} \left(\mathbf{X}, \lambda, x_0 \right) = \frac{1}{\sum_{i=1}^n w_i} \sum_{i=1}^n w_i \left( F_{\epsilon}\left(x_i \right) - F_{\mathcal{N}} \left(\psi^{\lambda, x_0} \left(x_i \right); \mu, \sigma \right) \right)^2
\end{equation}
```
